% Options for packages loaded elsewhere
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother


\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother


% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}



\setlength{\emergencystretch}{3em} % prevent overfull lines

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}



 


\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Privacy in an age of magical machines: A cross-cultural discovery of patterns of privacy in GenAI},
  pdfkeywords={data-capitalism, neocolonialism, genAI},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}


\title{Privacy in an age of magical machines: A cross-cultural discovery
of patterns of privacy in GenAI}
\author{Frederik J van Deventer}
\date{}
\begin{document}
\maketitle


\subsection{Introduction}\label{introduction}

Andreessen (2011) said: ``Software is eating the world'', but today it
might be more appropriate to say \emph{Data} is eating the world. There
is no longer a distinction to what can and cannot be captured in data
from nature (Walford 2018) to bodies (Van der Ploeg 2012). This
`datafication' (Cukier and Mayer-Schönberger 2014), tries to capture all
things into data, with data being a tradeable commodity (De Franceschi
and Lehmann 2015). Our daily lives are deeply connected with digital
technology that collects the what and when on our interactions as
capital (West 2019). The tools and distractions social media provides
turns us into consumer-workers (Puaschunder 2018) collecting data for
these companies whilst we enter into license agreements without being
able to control the data we produce (Thatcher, O'Sullivan, and Mahmoudi
2016) and approaching user agreements with a sense of apathy (Gómez
Ortega, Morales Ornelas, and Genç 2025).

Recent developments have seen generative Artificial Intelligence (genAI)
being available for the masses with Large Language Models (LLM's) having
a much easier interface for non-technical users. Often signified with
the moniker: AI or with magical stars (✨) it features in many
applications ranging from web search to (Google n.d.a; DuckDuckGo n.d.;
Microsoft n.d.a) personal communication such as Outlook ({``Chat with
{Copilot} in {Outlook} - {Microsoft Support}''} n.d.) or Whatsapp (Meta
n.d.). genAI is quickly becoming embedded in all sorts of business and
applications (Grantham-Philips 2024). Even more than in applications
beforehand users are encouraged to upload documents and provide more
context for the LLM to perform better.

GenAI has been lauded as tool that can help with productivity, can help
out in creative industries such as writing, music, film. But there is a
real darkside to it too. Besides the darkpatterns that are emerging and
the plausible, but false pictures and text it can produce, there are
many concerns about the energy abuse and infrastructure that is needed
for it to function. The impact it is having on learning and social
relations (Hou et al. 2025). Then there is the issue of bias by training
and the amplification of this bias (Lloyd 2018). Jailbreak prompts are a
way to subvert the main goals and to misuse LLM's (Perez and Ribeiro
2022). Which could lead to the extraction of personally identifiable
information (Li et al. 2023). In older models (GPT-2) training data
could be even be extracted after an attack (Carlini et al. 2021).

Concerning the use of personal data, this in itself should be alarming.
On top of that the terms of service for ChatGPT allow for OpenAI (the
company behind the popular chat based LLM) to make use of content
provided by users to improve services, and even to train models (OpenAi
n.d.). The same holds true for Anthropic (Claude) (Anthropic n.d.),
Google (Gemini) (Google n.d.b) and X (Grok) (X n.d.), even in areas with
stricter legal restrictions such as the European GDPR.

Finally there seems to be an inherent faith and push for Big Tech
companies to ``spread the good news'', to use religious language.

Recent news from Amazon, Microsoft and Google, the three largest
companies that provide infrastructure for software around the world,
mentions that they are investing in data centers in Uruguay (Google
2024), Thailand and Malaysia (Onag n.d.; Browne 2024; Chiang 2024),
Busan in South Korea (Microsoft 2024), Brazil (Microsoft, n.d.b) and
Chile (Amazon 2025; Cambero 2025). Who stands to benefit from these
transactions?

The extracting of data and the use of AI have become a powerful tool,
but at the same time data and AI ``entrench power assymetries and
engender new forms of structural violence and new inequities between the
Global South and North'' (Madianou 2024).

In recent years this phenomenon of data extraction has been described as
`data colonialism', as an exploitative transgression for profit.
(Couldry and Mejias 2019), digital colonialism (Kwet 2019) or `data
extractivism' (Brevini et al. 2024, pp 126), or in the case of African
nations even a `21st century scramble for africa' (Coleman 2018).

For this research however I would like to focus on data justice as
Jimenez further poses that a focus on justice instead of coloniality
would require `us to go further and actively attempt to eliminate the
injustices we observed' (Brevini et al. 2024, pp 133). Instead of only
looking at what that coloniality looks like and trying to prove this as
`colonial'.

As Kwet (2019) asks himself in the case of South Africa: `are cloud
centres built by Amazon, Microsoft, and Google good for the country'?
Software and the push for GenAI comes with its own agenda and is not
free from political or commercial agenda's (Kwet 2019). Abandoned
experiments as part of `AI for good' have been compared to leaving a
Lamborghini in the desert.

What then do we define as progress? Is it the mere evolution of
technological tools? Or the more petabytes we can store? Or is it
actually making a difference in peoples lives in a profound, meaningful
and positive way? And does importing more corporate ``Big Tech''
technology increase the way of life? Or does it further enrich Global
North pockets?

\subsection{Research Questions}\label{research-questions}

What actual gains can we see in productivity when we look at the use
genAI tools in the Global South and do they measure up to the costs made
towards that gain?

\subsubsection{Sub questions}\label{sub-questions}

\begin{itemize}
\tightlist
\item
  What are the costs of using genAI?
\item
  What uses are there?
\item
  How do costs differ over different types of use?
\item
  \emph{Who stands to gain?}
\end{itemize}

\subsection{Bonus questions}\label{bonus-questions}

\begin{quote}
Who owns the data?
\end{quote}

\begin{quote}
What influence can AI for \emph{Social Good} actually have? What is
\end{quote}

\begin{quote}
What is the relation of power and the created social standards In what
way is power big tech has exterted in and misused in policies concerning
data in lower income countries? - Lead with examples of policies in
powerful (Khanal, Zhang, and Taeihagh 2025)
\end{quote}

\begin{quote}
In what ways does GenAI express itself as a neocolonialist hegemonical
agent? - How do localized LLM's differ in bias towards policies from
commercial models widely available
\end{quote}

\begin{quote}
How is blind faith in techonological advancement towards ``Singularity''
influencing optimism towards creating an Artificial General
Intelligence? - forces of marketing and opportunities of use -
overpromises (Krokowski and Hirsch-Kreinsen 2025) -
\end{quote}

\begin{quote}
What local solutions can we offer for people groups who have a larger
gap towards digital accessibility
\end{quote}

\begin{quote}
\begin{itemize}
\tightlist
\item
  How do people use genAI cross-cultural study?
\item
  What patterns exist in the solution realm?
\item
  What intrinsic motvation
\end{itemize}
\end{quote}

\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-amazonWorksAWSSouth2025}
Amazon. 2025. {``In the Works -- {AWS South America} ({Chile}) {Region}
\textbar{} {AWS News Blog}.''}
https://aws.amazon.com/blogs/aws/coming-soon-aws-south-america-chile-region/.

\bibitem[\citeproctext]{ref-andreessenWhySoftwareEating2011}
Andreessen, Marc. 2011. {``Why {Software Is Eating} the {World}.''}
\emph{Andreessen Horowitz}.

\bibitem[\citeproctext]{ref-anthropicHowYouUse}
Anthropic. n.d. {``How {Do You Use Personal Data} in {Model Training}?
\textbar{} {Anthropic Privacy Center}.''}
https://privacy.claude.com/en/articles/10023555-how-do-you-use-personal-data-in-model-training.
Accessed November 17, 2025.

\bibitem[\citeproctext]{ref-brevini2024critiques}
Brevini, Benedetta, Irene Fubara-Manuel, Clément Le Ludec, Jakob Linaa
Jensen, Andrea Jimenez, and Jo Bates. 2024. {``Critiques of Data
Colonialism.''} In \emph{Dialogues in Data Power}, 120--37. Bristol
University Press.

\bibitem[\citeproctext]{ref-browneGoogleInvestux2412024}
Browne, Ryan. 2024. {``Google to Invest \$1 Billion in {Thailand} to
Build Data Center and Accelerate {AI} Growth.''} \emph{CNBC}.
https://www.cnbc.com/2024/09/30/google-to-invest-1-billion-in-thailand-data-center-and-ai-push.html.

\bibitem[\citeproctext]{ref-camberoAmazonSpendux2442025}
Cambero, Fabian. 2025. {``Amazon to Spend \$4 Billion on Cloud
Infrastructure in {Chile}.''} \emph{Reuters}, May.

\bibitem[\citeproctext]{ref-carliniExtractingTrainingData2021}
Carlini, Nicholas, Florian Tramer, Eric Wallace, Matthew Jagielski,
Ariel Herbert-Voss, Katherine Lee, Adam Roberts, et al. 2021.
{``Extracting {Training Data} from {Large Language Models}.''} arXiv.
\url{https://doi.org/10.48550/arXiv.2012.07805}.

\bibitem[\citeproctext]{ref-ChatCopilotOutlook}
{``Chat with {Copilot} in {Outlook} - {Microsoft Support}.''} n.d.
https://support.microsoft.com/en-us/topic/chat-with-copilot-in-outlook-8090e7b3-5b1d-4c6d-9b06-02edac062f58.
Accessed November 6, 2025.

\bibitem[\citeproctext]{ref-chiangMicrosoftOpenNew2024}
Chiang, Sheila. 2024. {``Microsoft to Open New Data Center in {Thailand}
as It Doubles down on {AI} and {Southeast Asia}.''} \emph{CNBC}.
https://www.cnbc.com/2024/05/02/microsoft-to-open-data-center-in-thailand-amid-southeast-asia-expansion.html.

\bibitem[\citeproctext]{ref-coleman2018digital}
Coleman, Danielle. 2018. {``Digital Colonialism: {The} 21st Century
Scramble for {Africa} Through the Extraction and Control of User Data
and the Limitations of Data Protection Laws.''} \emph{Mich. J. Race \&
L.} 24: 417.

\bibitem[\citeproctext]{ref-couldryDataColonialismRethinking2019}
Couldry, Nick, and Ulises A. Mejias. 2019. {``Data {Colonialism}:
{Rethinking Big Data}'s {Relation} to the {Contemporary Subject}.''}
\emph{Television \& New Media} 20 (4): 336--49.
\url{https://doi.org/10.1177/1527476418796632}.

\bibitem[\citeproctext]{ref-cukierRiseBigData2014}
Cukier, Kenneth, and Viktor Mayer-Schönberger. 2014. {``The {Rise} of
{Big Data}: {How It}'s {Changing} the {Way We Think} about the
{World}.''} In \emph{The {Best Writing} on {Mathematics} 2014}, edited
by Mircea Pitici, 20--32. Princeton University Press.
\url{https://doi.org/10.1515/9781400865307-003}.

\bibitem[\citeproctext]{ref-defranceschiDataTradeableCommodity2015}
De Franceschi, Alberto, and Michael Lehmann. 2015. {``Data as {Tradeable
Commodity} and {New Measures} for {Their Protection Essays}.''}
\emph{Italian Law Journal} 1 (1): 51--72.

\bibitem[\citeproctext]{ref-duckduckgoDuckaiDuckDuckGoHelp}
DuckDuckGo. n.d. {``Duck.ai - {DuckDuckGo Help Pages}.''}
https://duckduckgo.com/duckduckgo-help-pages/duckai. Accessed November
6, 2025.

\bibitem[\citeproctext]{ref-gomezortegaSurrenderingPowerlesnessGoverning2025}
Gómez Ortega, Alejandra, Hosana Morales Ornelas, and Uğur Genç. 2025.
{``Surrendering to {Powerlesness}: {Governing Personal Data Flows} in
{Generative AI}.''} In \emph{Proceedings of the 2025 {CHI Conference} on
{Human Factors} in {Computing Systems}}, 1--18. Yokohama Japan: ACM.
\url{https://doi.org/10.1145/3706598.3713504}.

\bibitem[\citeproctext]{ref-googleNewDataCenter2024}
Google. 2024. {``A New Data Center in {Latin America}.''} \emph{Google}.
https://blog.google/around-the-globe/google-latin-america/a-new-data-center-in-latin-america/.

\bibitem[\citeproctext]{ref-googleGoogleAIMode}
Google. n.d.a. {``Google {AI Mode} - a New Way to Search, Whatever's on
Your Mind.''} \emph{Google AI Mode - a New Way to Search, Whatever's on
Your Mind}. https://search.google/ways-to-search/ai-mode/. Accessed
November 6, 2025.

\bibitem[\citeproctext]{ref-googlePrivacyPolicyPrivacy}
Google. n.d.b. {``Privacy {Policy} -- {Privacy} \& {Terms} --
{Google}.''} https://policies.google.com/privacy. Accessed November 17,
2025.

\bibitem[\citeproctext]{ref-grantham-philipsAIBecomingIngrained2024}
Grantham-Philips, Wyatte. 2024. {``{AI} Is Becoming Ingrained in
Businesses Across Industries. {Where} Is It Going in 2025?''} \emph{AP
News}.
https://apnews.com/article/artificial-intelligence-interview-pwc-dan-priest-a0458061469735aed7af5fa49682e076.

\bibitem[\citeproctext]{ref-houAllRoadsLead2025}
Hou, Irene, Owen Man, Kate Hamilton, Srishty Muthusekaran, Jeffin
Johnykutty, Leili Zadeh, and Stephen MacNeil. 2025. {``'{All Roads Lead}
to {ChatGPT}': {How Generative AI} Is {Eroding Social Interactions} and
{Student Learning Communities}.''} In \emph{Proceedings of the 30th {ACM
Conference} on {Innovation} and {Technology} in {Computer Science
Education V}. 1}, 79--85. Nijmegen Netherlands: ACM.
\url{https://doi.org/10.1145/3724363.3729024}.

\bibitem[\citeproctext]{ref-khanalWhyHowPower2025}
Khanal, Shaleen, Hongzhou Zhang, and Araz Taeihagh. 2025. {``Why and How
Is the Power of {Big Tech} Increasing in the Policy Process? {The} Case
of Generative {AI}.''} \emph{Policy and Society} 44 (1): 52--69.
\url{https://doi.org/10.1093/polsoc/puae012}.

\bibitem[\citeproctext]{ref-krokowskiOncePromisedForever2025}
Krokowski, Thorben, and Hartmut Hirsch-Kreinsen. 2025. {``Once Promised,
Forever Believed? {An} Essay on the Blind Spots of {AI} Promises.''}
\emph{AI \& SOCIETY}, November.
\url{https://doi.org/10.1007/s00146-025-02765-1}.

\bibitem[\citeproctext]{ref-kwetDigitalColonialismUS2019}
Kwet, Michael. 2019. {``Digital Colonialism: {US} Empire and the New
Imperialism in the {Global South}.''} \emph{Race \& Class} 60 (4):
3--26. \url{https://doi.org/10.1177/0306396818823172}.

\bibitem[\citeproctext]{ref-liMultistepJailbreakingPrivacy2023}
Li, Haoran, Dadi Guo, Wei Fan, Mingshi Xu, Jie Huang, Fanpu Meng, and
Yangqiu Song. 2023. {``Multi-Step {Jailbreaking Privacy Attacks} on
{ChatGPT}.''} arXiv. \url{https://doi.org/10.48550/arXiv.2304.05197}.

\bibitem[\citeproctext]{ref-lloydBiasAmplificationArtificial2018}
Lloyd, Kirsten. 2018. {``Bias {Amplification} in {Artificial
Intelligence Systems}.''} arXiv.
\url{https://doi.org/10.48550/ARXIV.1809.07842}.

\bibitem[\citeproctext]{ref-madianou2024technocolonialism}
Madianou, Mirca. 2024. \emph{Technocolonialism: {When} Technology for
Good Is Harmful}. John Wiley \& Sons.

\bibitem[\citeproctext]{ref-metaMetaAIWhatsApp}
Meta. n.d. {``Meta {AI} in {WhatsApp}: {Chat}, {Create} \& {Get Things
Done}.''} \emph{WhatsApp.com}. https://www.whatsapp.com/meta-ai.
Accessed November 6, 2025.

\bibitem[\citeproctext]{ref-microsoftCelebratingCompletionNew2024}
Microsoft. 2024. {``Celebrating Completion of a New Datacenter in
{Busan}.''} \emph{Microsoft Local}.

\bibitem[\citeproctext]{ref-microsoftIntroducingNewBing}
Microsoft. n.d.a. {``Introducing the New {Bing}. {The AI-powered}
Assistant for Your Search.''}
https://www.microsoft.com/en-us/edge/features/the-new-bing. Accessed
November 6, 2025.

\bibitem[\citeproctext]{ref-microsoftMicrosoftDatacentersBrazil}
Microsoft. n.d.b. {``Microsoft Datacenters in {Brazil}.''}

\bibitem[\citeproctext]{ref-onagGoogleUnveilsux243B}
Onag, Gigi. n.d. {``Google Unveils \${3B} Cloud and Data Center
Investment in {Thailand} and {Malaysia}.''} \emph{Light Reading}.
https://www.lightreading.com/data-centers/google-unveils-3b-cloud-and-data-center-investment-in-thailand-and-malaysia.
Accessed December 8, 2025.

\bibitem[\citeproctext]{ref-openaiPrivacyPolicy}
OpenAi. n.d. {``Privacy Policy.''}
https://openai.com/policies/row-privacy-policy/. Accessed November 17,
2025.

\bibitem[\citeproctext]{ref-perezIgnorePreviousPrompt2022}
Perez, Fábio, and Ian Ribeiro. 2022. {``Ignore {Previous Prompt}:
{Attack Techniques For Language Models}.''} arXiv.
\url{https://doi.org/10.48550/arXiv.2211.09527}.

\bibitem[\citeproctext]{ref-puaschunder2018dignity}
Puaschunder, Julia M. 2018. {``Dignity and Utility of Privacy and
Information Sharing in the Digital Big Data Age.''} \emph{International
Journal of Commerce and Management Research} 5 (4): 62--70.

\bibitem[\citeproctext]{ref-thatcher2016data}
Thatcher, Jim, David O'Sullivan, and Dillon Mahmoudi. 2016. {``Data
Colonialism Through Accumulation by Dispossession: {New} Metaphors for
Daily Data.''} \emph{Environment and Planning D: Society and Space} 34
(6): 990--1006.

\bibitem[\citeproctext]{ref-van2012body}
Van der Ploeg, Irma. 2012. {``The Body as Data in the Age of
Information.''} In \emph{Routledge Handbook of Surveillance Studies},
176--83. Routledge.

\bibitem[\citeproctext]{ref-walford2018if}
Walford, Antonia. 2018. {``{`{If} Everything Is Information'}: Archives
and Collecting on the Frontiers of Data-Driven Science.''} In
\emph{Ethnography for a Data-Saturated World}, 105--27. Manchester
University Press.

\bibitem[\citeproctext]{ref-west2019data}
West, Sarah Myers. 2019. {``Data Capitalism: {Redefining} the Logics of
Surveillance and Privacy.''} \emph{Business \& Society} 58 (1): 20--41.

\bibitem[\citeproctext]{ref-xTermsServiceConsumer}
X. n.d. {``Terms of {Service} - {Consumer} \textbar{} {xAI}.''}
https://x.ai/legal/terms-of-service. Accessed November 17, 2025.

\end{CSLReferences}




\end{document}
