% Options for packages loaded elsewhere
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother


\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother


% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}



\setlength{\emergencystretch}{3em} % prevent overfull lines

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}



 


\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Avoiding the Turing Trap in an age of GenAI for increased human agency and freedom},
  pdfkeywords={human agency, agentic AI, CAIL, turing trap, AI
Literacy},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}


\title{Avoiding the Turing Trap in an age of GenAI for increased human
agency and freedom}
\author{Frederik J van Deventer}
\date{}
\begin{document}
\maketitle


\subsection{Introduction}\label{introduction}

Technological advances that herald and drive societal changes over the
last 12,000 years have only benefited ordinary people when ``landowning
and religious elites were not dominant enough to impose their vision and
extract all the surplus from the technologies'' (Johnson and Acemoglu
2023, 78).

We are again on the precipice of a big transformation for our society
with the wide adoption of Artificial Intelligence in all kinds of
sectors and embedded in all kinds of software. The public has never
adopted a technological platform so fast as it did with ChatGPT (Milmo
2023). Heralding the Generative Artificial Intelligence (genAI) boom
with the text-based models like ChatGPT, Claude and Gemini and
image-based products such as Dall-E and Midjourney being available for
the public.

Only a few years later we are slowly moving towards Agentic AI. Where
the Large Language Models (LLMs) that make the likes of Claude and
ChatGPT possible, are used to create (semi-)autonomous systems that
``reason'' and feedback into itself to perform the tasks or solutions
that it proposes. This marks a paradigm shift in the use of the web,
where not only humans shape its future (Yang et al. 2025). Because it is
so new rules, regulations or benchmarks are not yet in place and it is
``hard to distinguish genuine advances from hype'' (Kapoor et al. 2024).

Generative Artificial Intelligence (genAI), a specific form of
Artificial Intelligence harnessing the power of Large Language Models to
\emph{generate} new output based on vast and widely generalized data
sources, is rapidly transforming our lives. Ranging from reshaping how
we work (Joshi 2025) and how we learn (Perifanou and Economides 2025),
to changing criminal behaviour (Ferrara 2024) and even influencing
democratization (Cupać, Schopmans, and Tuncer-Ebetürk 2024).

With it there is this sense that we are closer to the holy grail of AI:
Artificial General Intelligence (AGI) or ``human-level intelligence'', a
``god-term'' and even ``devil-term'', where believers herald a utopia in
which AGI is able to solve everything, and on the other hand doomsayers
who believe it to become a catastrophe (Heaven 2025).

What we are seeing is a repetition of the technology industry
overpromising new techniques or technologies and packaging them as a
panacea for all kinds of problems or General Purpose Tools, while also
decreasing emphasis on a large variety of negative consequences that
this technology also brings. For example the profound physical impact
these new technologies are having on our environment (Crawford 2021,
49), the security issues of models with jailbreak attempts Carlini et
al. (2021), nudging political preferences(Cupać, Schopmans, and
Tuncer-Ebetürk 2024), reducing the quality of democracy decision-making
and the erosion of trust in legitimate institutions (Miklian and
Hoelscher, n.d.).

Myth and technology have a long history together. Collectively we adhere
to the idea that with technology and change we also converge to a better
reality, to progress, but believing this is a religious act itself
(Burdett 2014) with more religious overtones in what Winner (2018) calls
the ``Cult of Innovation''. We have seen similar rational and
``more-than-rational'' \emph{truths} with other technologies such as
blockchain (Gloerich 2025).

Technology is not merely a neutral phenomenon (Heyndels 2023; Morrow
2014) with inherently positive outcomes; there is a great need for
finding solutions that benefit us all.

\subsubsection{Gap}\label{gap}

We do not know how AI will reshape the future of our human development,
but there are some guesses we can make and historical lessons we can
take into account. In all the major jumps in technological advancement
that brought about societal change: steam engine, electricity,
communications, digital communication and now the more broadly available
statistical models that ``converse'' with us through chat like ChatGPT,
we have seen that regulation and policy fueled by ideology (both
benefiting the few or the many) has been the driver for the direction
this advancement would take us (Johnson and Acemoglu 2023, 57).

Much of the research regarding AI has been focussing on improving models
(Du, Yang, and Welleck 2025; Yao et al. 2025), effects on policy
(Khanal, Zhang, and Taeihagh 2025; Leslie and Perini 2024), impact on
education (Dai 2025; Han, Peng, and Liu 2025; Udeh 2025) on human
interactions (Hou et al. 2025), and how it can improve self-advocacy
(Register and Ko 2020) but not how it affects agency and freedom which
both count as indicative of improving human development (UNDP 2025).

Most of the AI Literacy literature agree that AI Literacy education
``empowers individuals to achieve humanistic outcomes'' (Pinski and
Benlian 2024).

\subsubsection{Hook}\label{hook}

\emph{``Technological development has invested the powers that be with
not only more efficient, better and more deadly instruments of coercion
but with the instruments of persuasion decidedly more efficient than
those hitherto used by the political bosses..''} - (Mathur, n.d.)

We need to figure a way out of what Brynjolfsson (2022) calls the
``Turing Trap'' where increasing technological and economic power create
a concentration of political power and find solutions, patterns,
systems, beyond policy where the benefits of transformational technology
like agentic AI are available to everyone.

Adolescents such as students in higher education institutes are often
already interacting with tools like ChatGPT (Dai 2025), so this study
primarily focuses on that group.

\subsection{Research Questions}\label{research-questions}

What kind of effects is GenAI having on human agency and freedom among
adolescents?

\subsubsection{Sub-questions}\label{sub-questions}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  How are students interacting with GenAI?
\item
  What kind of beliefs do students have about GenAI and its capabilities
  in relation to their own capabilities?
\item
  How is GenAI affecting the execution of tasks for studying?
\end{enumerate}

\subsection{Methodology}\label{methodology}

\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-brynjolfssonTuringTrapPromise2022}
Brynjolfsson, Erik. 2022. {``The {Turing Trap}: {The Promise} \& {Peril}
of {Human-Like Artificial Intelligence}.''} \emph{Daedalus} 151 (2):
272--87. \url{https://doi.org/10.1162/daed_a_01915}.

\bibitem[\citeproctext]{ref-burdett2014religion}
Burdett, Michael S. 2014. {``The Religion of Technology: {Transhumanism}
and the Myth of Progress.''} \emph{Religion and Transhumanism: The
Unknown Future of Human Enhancement}, 131.

\bibitem[\citeproctext]{ref-carliniExtractingTrainingData2021}
Carlini, Nicholas, Florian Tramer, Eric Wallace, Matthew Jagielski,
Ariel Herbert-Voss, Katherine Lee, Adam Roberts, et al. 2021.
{``Extracting {Training Data} from {Large Language Models}.''} arXiv.
\url{https://doi.org/10.48550/arXiv.2012.07805}.

\bibitem[\citeproctext]{ref-crawford2021atlas}
Crawford, Kate. 2021. \emph{The Atlas of {AI}: {Power}, Politics, and
the Planetary Costs of Artificial Intelligence}. Yale University Press.

\bibitem[\citeproctext]{ref-cupacDemocratizationAgeArtificial2024}
Cupać, Jelena, Hendrik Schopmans, and İrem Tuncer-Ebetürk. 2024.
{``Democratization in the Age of Artificial Intelligence: Introduction
to the Special Issue.''} \emph{Democratization} 31 (5): 899--921.
\url{https://doi.org/10.1080/13510347.2024.2338852}.

\bibitem[\citeproctext]{ref-daiWhyStudentsUse2025}
Dai, Yun. 2025. {``Why Students Use or Not Use Generative {AI}:
{Student} Conceptions, Concerns, and Implications for Engineering
Education.''} \emph{Digital Engineering} 4 (March): 100019.
\url{https://doi.org/10.1016/j.dte.2024.100019}.

\bibitem[\citeproctext]{ref-duOptimizingTemperatureLanguage2025}
Du, Weihua, Yiming Yang, and Sean Welleck. 2025. {``Optimizing
{Temperature} for {Language Models} with {Multi-Sample Inference}.''}
arXiv. \url{https://doi.org/10.48550/arXiv.2502.05234}.

\bibitem[\citeproctext]{ref-ferraraGenAIHumanityNefarious2024}
Ferrara, Emilio. 2024. {``{GenAI} Against Humanity: Nefarious
Applications of Generative Artificial Intelligence and Large Language
Models.''} \emph{Journal of Computational Social Science} 7 (1):
549--69. \url{https://doi.org/10.1007/s42001-024-00250-1}.

\bibitem[\citeproctext]{ref-gloerichReimaginingTruthMachine2025}
Gloerich, Inte. 2025. {``Reimagining the {Truth Machine}: {Blockchain
Imaginaries} Between the {Rational} and the {More-than-Rational}.''}
Dr., Utrecht University. \url{https://doi.org/10.33540/2726}.

\bibitem[\citeproctext]{ref-hanImpactGenAILearning2025}
Han, Xiaoli, Hongchao Peng, and Mingzhuo Liu. 2025. {``The Impact of
{GenAI} on Learning Outcomes: {A} Systematic Review and Meta-Analysis of
Experimental Studies.''} \emph{Educational Research Review} 48 (August):
100714. \url{https://doi.org/10.1016/j.edurev.2025.100714}.

\bibitem[\citeproctext]{ref-HeavenHowAGIBecamea}
Heaven, Will Douglas. 2025. {``How {AGI} Became the Most Consequential
Conspiracy Theory of Our Time.''} \emph{MIT Technology Review}.
https://www.technologyreview.com/2025/10/30/1127057/agi-conspiracy-theory-artifcial-general-intelligence/.

\bibitem[\citeproctext]{ref-heyndelsTechnologyNeutrality2023}
Heyndels, Sybren. 2023. {``Technology and {Neutrality}.''}
\emph{Philosophy \& Technology} 36 (4): 75.
\url{https://doi.org/10.1007/s13347-023-00672-1}.

\bibitem[\citeproctext]{ref-houAllRoadsLead2025}
Hou, Irene, Owen Man, Kate Hamilton, Srishty Muthusekaran, Jeffin
Johnykutty, Leili Zadeh, and Stephen MacNeil. 2025. {``'{All Roads Lead}
to {ChatGPT}': {How Generative AI} Is {Eroding Social Interactions} and
{Student Learning Communities}.''} In \emph{Proceedings of the 30th {ACM
Conference} on {Innovation} and {Technology} in {Computer Science
Education V}. 1}, 79--85. Nijmegen Netherlands: ACM.
\url{https://doi.org/10.1145/3724363.3729024}.

\bibitem[\citeproctext]{ref-johnsonPowerProgressOur2023}
Johnson, Simon, and Daron Acemoglu. 2023. \emph{Power and {Progress}:
{Our Thousand-Year Struggle Over Technology} and {Prosperity}}. Hachette
UK.

\bibitem[\citeproctext]{ref-joshi2025transformative}
Joshi, Satyadhar. 2025. {``The Transformative Role of Agentic {GenAI} in
Shaping Workforce Development and Education in the {US}.''}
\emph{Available at SSRN 5133376}.

\bibitem[\citeproctext]{ref-kapoorAIAgentsThat2024}
Kapoor, Sayash, Benedikt Stroebl, Zachary S. Siegel, Nitya Nadgir, and
Arvind Narayanan. 2024. {``{AI Agents That Matter}.''} arXiv.
\url{https://doi.org/10.48550/ARXIV.2407.01502}.

\bibitem[\citeproctext]{ref-khanalWhyHowPower2025}
Khanal, Shaleen, Hongzhou Zhang, and Araz Taeihagh. 2025. {``Why and How
Is the Power of {Big Tech} Increasing in the Policy Process? {The} Case
of Generative {AI}.''} \emph{Policy and Society} 44 (1): 52--69.
\url{https://doi.org/10.1093/polsoc/puae012}.

\bibitem[\citeproctext]{ref-leslieFutureShockGenerative2024}
Leslie, David, and Antonella Maia Perini. 2024. {``Future {Shock}:
{Generative AI} and the {International AI Policy} and {Governance
Crisis}.''} \emph{Harvard Data Science Review}, no. Special Issue 5
(May). \url{https://doi.org/10.1162/99608f92.88b4cc98}.

\bibitem[\citeproctext]{ref-liMultistepJailbreakingPrivacy2023}
Li, Haoran, Dadi Guo, Wei Fan, Mingshi Xu, Jie Huang, Fanpu Meng, and
Yangqiu Song. 2023. {``Multi-Step {Jailbreaking Privacy Attacks} on
{ChatGPT}.''} arXiv. \url{https://doi.org/10.48550/arXiv.2304.05197}.

\bibitem[\citeproctext]{ref-mathurTechnologyFreedom}
Mathur, A B. n.d. {``Technology and {Freedom}.''}

\bibitem[\citeproctext]{ref-miklianNewDigitalDivide}
Miklian, Jason, and Kristian Hoelscher. n.d. {``A New Digital Divide?
{Coder} Worldviews, the {`{Slop} Economy,'} and Democracy in the Age of
{AI}.''}

\bibitem[\citeproctext]{ref-milmoChatGPTReaches1002023}
Milmo, Dan. 2023. {``{ChatGPT} Reaches 100 Million Users Two Months
After Launch.''} \emph{The Guardian}, February.

\bibitem[\citeproctext]{ref-morrowWhenTechnologiesMakes2014}
Morrow, David R. 2014. {``When {Technologies Makes Good People Do Bad
Things}: {Another Argument Against} the {Value-Neutrality} of
{Technologies}.''} \emph{Science and Engineering Ethics} 20 (2):
329--43. \url{https://doi.org/10.1007/s11948-013-9464-1}.

\bibitem[\citeproctext]{ref-perezIgnorePreviousPrompt2022}
Perez, Fábio, and Ian Ribeiro. 2022. {``Ignore {Previous Prompt}:
{Attack Techniques For Language Models}.''} arXiv.
\url{https://doi.org/10.48550/arXiv.2211.09527}.

\bibitem[\citeproctext]{ref-perifanouCollaborativeUsesGenAI2025}
Perifanou, Maria, and Anastasios A. Economides. 2025. {``Collaborative
{Uses} of {GenAI Tools} in {Project-Based Learning}.''} \emph{Education
Sciences} 15 (3): 354. \url{https://doi.org/10.3390/educsci15030354}.

\bibitem[\citeproctext]{ref-pinskiAILiteracyUsers2024}
Pinski, Marc, and Alexander Benlian. 2024. {``{AI} Literacy for Users --
{A} Comprehensive Review and Future Research Directions of Learning
Methods, Components, and Effects.''} \emph{Computers in Human Behavior:
Artificial Humans} 2 (1): 100062.
\url{https://doi.org/10.1016/j.chbah.2024.100062}.

\bibitem[\citeproctext]{ref-registerLearningMachineLearning2020}
Register, Yim, and Amy J. Ko. 2020. {``Learning {Machine Learning} with
{Personal Data Helps Stakeholders Ground Advocacy Arguments} in {Model
Mechanics}.''} In \emph{Proceedings of the 2020 {ACM Conference} on
{International Computing Education Research}}, 67--78. Virtual Event New
Zealand: ACM. \url{https://doi.org/10.1145/3372782.3406252}.

\bibitem[\citeproctext]{ref-udehRoleGenerativeAI2025}
Udeh, Chinemelum Goodness. 2025. {``The Role of Generative {AI} in
Personalized Learning for Higher Education.''} \emph{World Journal of
Advanced Engineering Technology and Sciences} 14 (2): 205--7.
\url{https://doi.org/10.30574/wjaets.2025.14.2.0077}.

\bibitem[\citeproctext]{ref-undpHumanDevelopmentReport2025}
UNDP. 2025. {``Human Development Report 2025.''} \emph{UNDP (United
Nations Development Programme)}.

\bibitem[\citeproctext]{ref-winnerCultInnovationIts2018}
Winner, Langdon. 2018. {``The {Cult} of {Innovation}: {Its Myths} and
{Rituals}.''} In \emph{Engineering a {Better Future}: {Interplay}
Between {Engineering}, {Social Sciences}, and {Innovation}}, edited by
Eswaran Subrahmanian, Toluwalogo Odumosu, and Jeffrey Y. Tsao, 61--73.
Cham: Springer International Publishing.
\url{https://doi.org/10.1007/978-3-319-91134-2_8}.

\bibitem[\citeproctext]{ref-yangAgenticWebWeaving2025}
Yang, Yingxuan, Mulei Ma, Yuxuan Huang, Huacan Chai, Chenyu Gong, Haoran
Geng, Yuanjian Zhou, et al. 2025. {``Agentic {Web}: {Weaving} the {Next
Web} with {AI Agents}.''} arXiv.
\url{https://doi.org/10.48550/arXiv.2507.21206}.

\bibitem[\citeproctext]{ref-yaoCAReDiOEnhancingCultural2025}
Yao, Jing, Xiaoyuan Yi, Jindong Wang, Zhicheng Dou, and Xing Xie. 2025.
{``{CAReDiO}: {Enhancing Cultural Alignment} of {LLM} via
{Representativeness} and {Distinctiveness Guided Data Optimization},''}
October.

\end{CSLReferences}




\end{document}
