% Options for packages loaded elsewhere
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother


\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother


% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}



\setlength{\emergencystretch}{3em} % prevent overfull lines

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}



 


\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Putting human agency at the forefront in agentic AI},
  pdfkeywords={human agency, agentic AI, CAIL, turing trap, AI Literacy,
belief},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}


\title{Putting human agency at the forefront in agentic AI}
\author{Frederik J van Deventer}
\date{}
\begin{document}
\maketitle


\subsection{Introduction}\label{introduction}

A shift in world order, a challenge with climate change, and more potent
ways to utilize our data with a resurgence of Artificial Intellegince
(AI) in the last few years have us all on the edge of our seat. We are
no longer sure what is real or what is generated {[}26{]}, we are
promised that AI will solve climate change {[}37{]} while at the same
time we are seeing an ``AI Gold Rush'' where more and more electrical
power use is being needed to sustain the growth of AI usage {[}23{]}.

The permeation of AI in our daily digital lives is almost inescapable.
The biggest communication platforms and work applications have an AI
function embedded, for example Copilot throughout Microsoft Office
Suite, Meta AI in WhatsApp and Google Search giving Gemini generated
results as top result. Signified with ``AI'' or ✨ it is usually a form
of a Large Language Model (LLM) that interprets requests and gives
human-like answers. With a new direction called \emph{Agentic AI} these
LLMs (LLMs) that make the likes of Claude and ChatGPT possible, are used
to create (semi-)autonomous systems that ``reason'' and feedback into
itself to perform the tasks or solutions that it proposes. This marks a
paradigm shift in the use of the web, where not only humans shape its
future {[}38{]}. Because it is so new rules, regulations or benchmarks
are not yet in place and it is ``hard to distinguish genuine advances
from hype'' {[}18{]}. The advances made in AI are hoped to work towards
a goal of Artifical General Intelligence (AGI), bringing ``human-like
intelligence'' {[}1{]}.

Talk of being ``falling behind'' {[}12{]} or a ``skill gap'' {[}5, 36{]}
or ``losing a race'' {[}10{]} fuels anxiety of missing out {[}4, 40{]}.
If AI automates or replaces human labor instead of augmenting it, it
creates the risk of concentrating wealth and power, the \emph{Turing
Trap} {[}2{]}.

Besides the future of labour there are other real and pressing issues
for us to contend with. For example the profound physical impact these
new technologies are having on our environment {[}6{]}, the security
issues of models with jailbreak attempts {[}3{]}, nudging political
preferences{[}7{]}, reducing the quality of democracy decision-making
and the erosion of trust in legitimate institutions {[}25{]}.

Young people around the globe are entering a different digital reality
than their parents, where AI is playing a big part. Much of the
adolescent student population is already using tools like ChatGPT
{[}8{]}. How then do we avoid the Turing Trap, and increase agency for
(young) humans instead of infringing upon it?

\subsection{Gap}\label{gap}

Much of the research regarding AI has been focussing on improving models
{[}11, 39{]}, improved productivity in writing code {[}20{]}, effects on
policy {[}19, 21{]}, impact on education {[}8, 14, 35{]} on human
interactions {[}16{]}, and how it can improve self-advocacy {[}33{]}.

In the course of history, human development has been argued to increase
where agency and liberty increase {[}32, 34{]}. Agentic AI and AI goals
in general, largely seem to focus on increasing abilities for models or
replacing tedious tasks performed by humans. While research has been
done about future projections of human agency in decision making
{[}30{]} and in education {[}28{]} it does not deal with how to increase
human agency for humans now.

AI literacy is one field where demystification and explanation are seen
to increase people's insight in systems {[}31{]} and help to reduce
anthropomorphism {[}9{]} which in itself can lead to false perceptions
of AI abilities {[}13{]}.

\subsection{Hook}\label{hook}

\emph{``Technological development has invested the powers that be with
not only more efficient, better and more deadly instruments of coercion
but with the instruments of persuasion decidedly more efficient than
those hitherto used by the political bosses..''} - {[}24{]}

If we want to create circumstances for technology where humans can
thrive we need to be able to accept a pluralistic view of technology. As
technology is never just neutral nor is it inherently good or bad {[}15,
27{]}. Regulation and policy fueled by ideology (both benefiting the few
or the many) has been the driver for the direction this advancement
would take us {[}17{]}.

\subsection{Research Questions}\label{research-questions}

How can adolescents achieve increased human agency and freedom in a
world with agentic AI?

\subsubsection{Sub-questions}\label{sub-questions}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  How are students interacting with agentic AI or GenAI?
\item
  How is AI affecting the execution of day-to-day tasks (e.g.~study,
  work, household chores)?
\item
  What beliefs do adolescents have about AI and its capabilities in
  relation to their own capabilities?
\item
  In what ways are AI Literacies affecting these beliefs?
\item
  How do adolescents discern hype from reality?
\item
  In what ways can understanding of these phenomena enhance their
  freedom of choice?
\end{enumerate}

\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{0}{0}
\bibitem[\citeproctext]{ref-blili-hamelinStopTreatingAGI2025}
\CSLLeftMargin{1. }%
\CSLRightInline{Blili-Hamelin, B. et al.: Stop treating {``{AGI}''} as
the north-star goal of {AI} research, (2025).
https://doi.org/\href{https://doi.org/10.48550/ARXIV.2502.03689}{10.48550/ARXIV.2502.03689}.}

\bibitem[\citeproctext]{ref-brynjolfssonTuringTrapPromise2022}
\CSLLeftMargin{2. }%
\CSLRightInline{Brynjolfsson, E.: The {Turing Trap}: {The Promise} \&
{Peril} of {Human-Like Artificial Intelligence}. Daedalus. 151, 2,
272--287 (2022).
https://doi.org/\href{https://doi.org/10.1162/daed_a_01915}{10.1162/daed\_a\_01915}.}

\bibitem[\citeproctext]{ref-carliniExtractingTrainingData2021}
\CSLLeftMargin{3. }%
\CSLRightInline{Carlini, N. et al.: Extracting {Training Data} from
{Large Language Models}, \url{https://arxiv.org/abs/2012.07805}, (2021).
https://doi.org/\href{https://doi.org/10.48550/arXiv.2012.07805}{10.48550/arXiv.2012.07805}.}

\bibitem[\citeproctext]{ref-chen2025exploring}
\CSLLeftMargin{4. }%
\CSLRightInline{Chen, Y.H.: Exploring the influence of generative
artificial intelligence anxiety on problematic {ChatGPT} use: The fear
of missing out and personality traits. Universal Access in the
Information Society. 24, 3, 2477--2490 (2025).}

\bibitem[\citeproctext]{ref-chrisingerSolutionLiesEducation2019}
\CSLLeftMargin{5. }%
\CSLRightInline{Chrisinger, D.: The solution lies in education:
Artificial intelligence \& the skills gap. On the Horizon: The
International Journal of Learning Futures. 27, 1, 1--4 (2019).
https://doi.org/\href{https://doi.org/10.1108/OTH-03-2019-096}{10.1108/OTH-03-2019-096}.}

\bibitem[\citeproctext]{ref-crawford2021atlas}
\CSLLeftMargin{6. }%
\CSLRightInline{Crawford, K.: The atlas of {AI}: {Power}, politics, and
the planetary costs of artificial intelligence. Yale University Press
(2021).}

\bibitem[\citeproctext]{ref-cupacDemocratizationAgeArtificial2024}
\CSLLeftMargin{7. }%
\CSLRightInline{Cupać, J. et al.: Democratization in the age of
artificial intelligence: Introduction to the special issue.
Democratization. 31, 5, 899--921 (2024).
https://doi.org/\href{https://doi.org/10.1080/13510347.2024.2338852}{10.1080/13510347.2024.2338852}.}

\bibitem[\citeproctext]{ref-daiWhyStudentsUse2025}
\CSLLeftMargin{8. }%
\CSLRightInline{Dai, Y.: Why students use or not use generative {AI}:
{Student} conceptions, concerns, and implications for engineering
education. Digital Engineering. 4, 100019 (2025).
https://doi.org/\href{https://doi.org/10.1016/j.dte.2024.100019}{10.1016/j.dte.2024.100019}.}

\bibitem[\citeproctext]{ref-drugaHowChildrensPerceptions2021}
\CSLLeftMargin{9. }%
\CSLRightInline{Druga, S., Ko, A.J.: How do children's perceptions of
machine intelligence change when training and coding smart programs? In:
Interaction {Design} and {Children}. pp. 49--61 ACM, Athens Greece
(2021).
https://doi.org/\href{https://doi.org/10.1145/3459990.3460712}{10.1145/3459990.3460712}.}

\bibitem[\citeproctext]{ref-du2025you}
\CSLLeftMargin{10. }%
\CSLRightInline{du Toit, F.: You are not falling behind, but you could
lose the race. MoneyMarketing. 2025, 6, 13--13 (2025).}

\bibitem[\citeproctext]{ref-duOptimizingTemperatureLanguage2025}
\CSLLeftMargin{11. }%
\CSLRightInline{Du, W. et al.: Optimizing {Temperature} for {Language
Models} with {Multi-Sample Inference},
\url{https://arxiv.org/abs/2502.05234}, (2025).
https://doi.org/\href{https://doi.org/10.48550/arXiv.2502.05234}{10.48550/arXiv.2502.05234}.}

\bibitem[\citeproctext]{ref-ederFallingAdoptionCurve2025}
\CSLLeftMargin{12. }%
\CSLRightInline{Eder, M., Sjøvaag, H.: Falling behind the adoption
curve: {Local} journalism's struggle for innovation in the {AI}
transformation. Journal of Media Business Studies. 22, 4, 325--343
(2025).
https://doi.org/\href{https://doi.org/10.1080/16522354.2025.2473301}{10.1080/16522354.2025.2473301}.}

\bibitem[\citeproctext]{ref-guestUncriticalAdoptionAI2025}
\CSLLeftMargin{13. }%
\CSLRightInline{Guest, O. et al.: Against the {Uncritical Adoption} of
'{AI}' {Technologies} in {Academia}, (2025).
https://doi.org/\href{https://doi.org/10.5281/ZENODO.17065099}{10.5281/ZENODO.17065099}.}

\bibitem[\citeproctext]{ref-hanImpactGenAILearning2025}
\CSLLeftMargin{14. }%
\CSLRightInline{Han, X. et al.: The impact of {GenAI} on learning
outcomes: {A} systematic review and meta-analysis of experimental
studies. Educational Research Review. 48, 100714 (2025).
https://doi.org/\href{https://doi.org/10.1016/j.edurev.2025.100714}{10.1016/j.edurev.2025.100714}.}

\bibitem[\citeproctext]{ref-heyndelsTechnologyNeutrality2023}
\CSLLeftMargin{15. }%
\CSLRightInline{Heyndels, S.: Technology and {Neutrality}. Philosophy \&
Technology. 36, 4, 75 (2023).
https://doi.org/\href{https://doi.org/10.1007/s13347-023-00672-1}{10.1007/s13347-023-00672-1}.}

\bibitem[\citeproctext]{ref-houAllRoadsLead2025}
\CSLLeftMargin{16. }%
\CSLRightInline{Hou, I. et al.: '{All Roads Lead} to {ChatGPT}': {How
Generative AI} is {Eroding Social Interactions} and {Student Learning
Communities}. In: Proceedings of the 30th {ACM Conference} on
{Innovation} and {Technology} in {Computer Science Education V}. 1. pp.
79--85 ACM, Nijmegen Netherlands (2025).
https://doi.org/\href{https://doi.org/10.1145/3724363.3729024}{10.1145/3724363.3729024}.}

\bibitem[\citeproctext]{ref-johnsonPowerProgressOur2023}
\CSLLeftMargin{17. }%
\CSLRightInline{Johnson, S., Acemoglu, D.: Power and {Progress}: {Our
Thousand-Year Struggle Over Technology} and {Prosperity}. Hachette UK
(2023).}

\bibitem[\citeproctext]{ref-kapoorAIAgentsThat2024}
\CSLLeftMargin{18. }%
\CSLRightInline{Kapoor, S. et al.: {AI Agents That Matter}, (2024).
https://doi.org/\href{https://doi.org/10.48550/ARXIV.2407.01502}{10.48550/ARXIV.2407.01502}.}

\bibitem[\citeproctext]{ref-khanalWhyHowPower2025}
\CSLLeftMargin{19. }%
\CSLRightInline{Khanal, S. et al.: Why and how is the power of {Big
Tech} increasing in the policy process? {The} case of generative {AI}.
Policy and Society. 44, 1, 52--69 (2025).
https://doi.org/\href{https://doi.org/10.1093/polsoc/puae012}{10.1093/polsoc/puae012}.}

\bibitem[\citeproctext]{ref-kumarIntuitionEvidenceMeasuring2025}
\CSLLeftMargin{20. }%
\CSLRightInline{Kumar, A. et al.: Intuition to {Evidence}: {Measuring
AI}'s {True Impact} on {Developer Productivity},
\url{https://arxiv.org/abs/2509.19708}, (2025).
https://doi.org/\href{https://doi.org/10.48550/arXiv.2509.19708}{10.48550/arXiv.2509.19708}.}

\bibitem[\citeproctext]{ref-leslieFutureShockGenerative2024}
\CSLLeftMargin{21. }%
\CSLRightInline{Leslie, D., Perini, A.M.: Future {Shock}: {Generative
AI} and the {International AI Policy} and {Governance Crisis}. Harvard
Data Science Review. Special Issue 5, (2024).
https://doi.org/\href{https://doi.org/10.1162/99608f92.88b4cc98}{10.1162/99608f92.88b4cc98}.}

\bibitem[\citeproctext]{ref-liMultistepJailbreakingPrivacy2023}
\CSLLeftMargin{22. }%
\CSLRightInline{Li, H. et al.: Multi-step {Jailbreaking Privacy Attacks}
on {ChatGPT}, \url{https://arxiv.org/abs/2304.05197}, (2023).
https://doi.org/\href{https://doi.org/10.48550/arXiv.2304.05197}{10.48550/arXiv.2304.05197}.}

\bibitem[\citeproctext]{ref-linExplodingAIPower2024}
\CSLLeftMargin{23. }%
\CSLRightInline{Lin, L. et al.: Exploding {AI Power Use}: An
{Opportunity} to {Rethink Grid Planning} and {Management}. In: The 15th
{ACM International Conference} on {Future} and {Sustainable Energy
Systems}. pp. 434--441 ACM, Singapore Singapore (2024).
https://doi.org/\href{https://doi.org/10.1145/3632775.3661959}{10.1145/3632775.3661959}.}

\bibitem[\citeproctext]{ref-mathurTechnologyFreedom}
\CSLLeftMargin{24. }%
\CSLRightInline{Mathur, A.B.: Technology and {Freedom}.}

\bibitem[\citeproctext]{ref-miklianNewDigitalDivide}
\CSLLeftMargin{25. }%
\CSLRightInline{Miklian, J., Hoelscher, K.: A new digital divide?
{Coder} worldviews, the {``{Slop} economy,''} and democracy in the age
of {AI}.}

\bibitem[\citeproctext]{ref-millerAIHyperrealismWhy2023}
\CSLLeftMargin{26. }%
\CSLRightInline{Miller, E.J. et al.: {AI Hyperrealism}: {Why AI Faces
Are Perceived} as {More Real Than Human Ones}. Psychological Science.
34, 12, 1390--1403 (2023).
https://doi.org/\href{https://doi.org/10.1177/09567976231207095}{10.1177/09567976231207095}.}

\bibitem[\citeproctext]{ref-morrowWhenTechnologiesMakes2014}
\CSLLeftMargin{27. }%
\CSLRightInline{Morrow, D.R.: When {Technologies Makes Good People Do
Bad Things}: {Another Argument Against} the {Value-Neutrality} of
{Technologies}. Science and Engineering Ethics. 20, 2, 329--343 (2014).
https://doi.org/\href{https://doi.org/10.1007/s11948-013-9464-1}{10.1007/s11948-013-9464-1}.}

\bibitem[\citeproctext]{ref-moutaWhereAgencyMoving2025}
\CSLLeftMargin{28. }%
\CSLRightInline{Mouta, A. et al.: {``{Where} is {Agency Moving} to?''}:
{Exploring} the {Interplay} between {AI Technologies} in {Education} and
{Human Agency}. Digital Society. 4, 2, 49 (2025).
https://doi.org/\href{https://doi.org/10.1007/s44206-025-00203-9}{10.1007/s44206-025-00203-9}.}

\bibitem[\citeproctext]{ref-perezIgnorePreviousPrompt2022}
\CSLLeftMargin{29. }%
\CSLRightInline{Perez, F., Ribeiro, I.: Ignore {Previous Prompt}:
{Attack Techniques For Language Models},
\url{https://arxiv.org/abs/2211.09527}, (2022).
https://doi.org/\href{https://doi.org/10.48550/arXiv.2211.09527}{10.48550/arXiv.2211.09527}.}

\bibitem[\citeproctext]{ref-pewFutureHuman2023}
\CSLLeftMargin{30. }%
\CSLRightInline{Pew Research Center et al.: The {Future} of {Human
Agency}. (2023).}

\bibitem[\citeproctext]{ref-pinskiAILiteracyUsers2024}
\CSLLeftMargin{31. }%
\CSLRightInline{Pinski, M., Benlian, A.: {AI} literacy for users -- {A}
comprehensive review and future research directions of learning methods,
components, and effects. Computers in Human Behavior: Artificial Humans.
2, 1, 100062 (2024).
https://doi.org/\href{https://doi.org/10.1016/j.chbah.2024.100062}{10.1016/j.chbah.2024.100062}.}

\bibitem[\citeproctext]{ref-PradosdelaEscosura_2022}
\CSLLeftMargin{32. }%
\CSLRightInline{Prados de la Escosura, L.: Human development and the
path to freedom: 1870 to the present. Cambridge University Press,
Cambridge (2022).}

\bibitem[\citeproctext]{ref-registerLearningMachineLearning2020}
\CSLLeftMargin{33. }%
\CSLRightInline{Register, Y., Ko, A.J.: Learning {Machine Learning} with
{Personal Data Helps Stakeholders Ground Advocacy Arguments} in {Model
Mechanics}. In: Proceedings of the 2020 {ACM Conference} on
{International Computing Education Research}. pp. 67--78 ACM, Virtual
Event New Zealand (2020).
https://doi.org/\href{https://doi.org/10.1145/3372782.3406252}{10.1145/3372782.3406252}.}

\bibitem[\citeproctext]{ref-senDevelopmentFreedom1999}
\CSLLeftMargin{34. }%
\CSLRightInline{Sen, A.: Development as {Freedom}. Oxford University
Press (1999).}

\bibitem[\citeproctext]{ref-udehRoleGenerativeAI2025}
\CSLLeftMargin{35. }%
\CSLRightInline{Udeh, C.G.: The role of generative {AI} in personalized
learning for higher education. World Journal of Advanced Engineering
Technology and Sciences. 14, 2, 205--207 (2025).
https://doi.org/\href{https://doi.org/10.30574/wjaets.2025.14.2.0077}{10.30574/wjaets.2025.14.2.0077}.}

\bibitem[\citeproctext]{ref-vanhaeftenBridgingAISkills2024}
\CSLLeftMargin{36. }%
\CSLRightInline{Van Haeften, W. et al.: Bridging the {AI Skills Gap} in
{Europe}: {A Detailed Analysis} of {AI Skills} and {Roles}. In:
Resilience {Through Digital Innovation}: {Enabling} the {Twin
Transition}. pp. 385--402 University of Maribor Press (2024).
https://doi.org/\href{https://doi.org/10.18690/um.fov.4.2024.22}{10.18690/um.fov.4.2024.22}.}

\bibitem[\citeproctext]{ref-PrioritisingAIClimate2024}
\CSLLeftMargin{37. }%
\CSLRightInline{Woodcock, N.: Prioritising {AI} over climate change
would be catastrophic, (2024).}

\bibitem[\citeproctext]{ref-yangAgenticWebWeaving2025}
\CSLLeftMargin{38. }%
\CSLRightInline{Yang, Y. et al.: Agentic {Web}: {Weaving} the {Next Web}
with {AI Agents}, \url{https://arxiv.org/abs/2507.21206}, (2025).
https://doi.org/\href{https://doi.org/10.48550/arXiv.2507.21206}{10.48550/arXiv.2507.21206}.}

\bibitem[\citeproctext]{ref-yaoCAReDiOEnhancingCultural2025}
\CSLLeftMargin{39. }%
\CSLRightInline{Yao, J. et al.: {CAReDiO}: {Enhancing Cultural
Alignment} of {LLM} via {Representativeness} and {Distinctiveness Guided
Data Optimization}. (2025).}

\bibitem[\citeproctext]{ref-yuFearMissingOut2025}
\CSLLeftMargin{40. }%
\CSLRightInline{Yu, C.-L.: Fear of missing out on {AI}: {A}
psychological cost of technological revolution. Telematics and
Informatics Reports. 20, 100277 (2025).
https://doi.org/\href{https://doi.org/10.1016/j.teler.2025.100277}{10.1016/j.teler.2025.100277}.}

\end{CSLReferences}




\end{document}
